\chapter{Conclusion}
This report highlights the need for further research aiming at the achievement of faster and more capable computer systems. This research can occur in multiple research directions and may involve several improvements in multiple components of the \acrshort{CPU} in order to solve the existing performance bottlenecks.

One of the most relevant problems is related to the manipulation of data utilized by the processor, since it can have a big impact in some modern work fields like Machine Learning, AI, and image and video processing. Leveraging the benefits of advanced memory manipulation strategies like data prefetching, data streaming or vector/SIMD operations is going to be essential to provide \acrfull{GPP} with the improved performance and efficiency that modern workloads and applications require.

The proposed work to be developed during this thesis focuses on implementing and evaluating the performance of the novelty solution of UVE with Data Streaming Support on the CVA6 6-stage CPU. The first step will be to experiment and define the implementation of the proposed streaming engine on the CPU pipeline with the added modifications to all associated functional units and required signals and data structures. After finalizing the hardware description of this modified CPU, a series of tests will be conducted to evaluate several characteristics of the system, like performance, efficiency, and footprint. 

At the end of this thesis, the work developed is expected to provide a new set of tools for further research and improvements on UVE with Data Streaming Support.








\begin{landscape}

\begin{figure}

\section{Gantt Chart}
    
   
\definecolor{groupblue}{RGB}{0, 157, 224}
\definecolor{barblue}{RGB}{173, 216, 230}
\resizebox{1.55\textwidth}{!}{
\begin{ganttchart}[vgrid={draw=none, dotted},%
            %today=15,%
            %today offset=.5,%
            %today label=Heute,%
            %progress=today,%
            y unit title=0.7cm,
            y unit chart=0.6cm,
            bar incomplete/.append style={fill=red},%
            progress label text=  {\quad\pgfmathprintnumber[precision=0,verbatim]{#1}\%}%
            ]{1}{28}
\gantttitle[]{Novel RISC-V Extension for Data Streaming Complex Patterns}{28}\\
\gantttitle[]{2023}{8}                
    \gantttitle[]{2024}{20}\\
    \gantttitle{Sep}{2}                    
    \gantttitle{Oct}{2}
    \gantttitle{Nov}{2}
    \gantttitle{Dec}{2}
    \gantttitle{Jan}{2}
    \gantttitle{Feb}{2}
    \gantttitle{Mar}{2}
    \gantttitle{Apr}{2}
    \gantttitle{May}{2}
    \gantttitle{Jun}{2}
    \gantttitle{Jul}{2}
    \gantttitle{Aug}{2}
    \gantttitle{Sep}{2}
    \gantttitle{Oct}{2} \\
  \ganttbar[bar/.append style={fill=barblue}]{Work enviroment configuration}{1}{7} \\
  
  \ganttbar[bar/.append style={fill=barblue}]{Writting the report}{6}{9} \\
%%%%%%%%%%%%%%%%%% 2024 %%%%%%%%%%%%%%%%%%%%%%
  
  \ganttbar[bar/.append style={fill=groupblue}] {\textbf{Proposed Architeture Definition and Consolidation}}{11}{13} \\

  \ganttbar[bar/.append style={fill=barblue}] {Analysis of the CVA6 memory interface}{11}{13} \\
  
  
  \ganttbar[bar/.append style={fill=groupblue}]{\textbf{Modification of CVA6 for UVE support with data streaming engine}}{11}{17} \\
  \ganttbar[bar/.append style={fill=barblue}]{Extension of deconding units}{5}{14} \\
  \ganttbar[bar/.append style={fill=barblue}]{Implementation of streaming engine units}{14}{17} \\

  
  \ganttbar[bar/.append style={fill=groupblue}] {\textbf{Design of workload test for the modified CVA6}}{17}{18} \\
  
  \ganttbar[bar/.append style={fill=groupblue}] {\textbf{Execution of tests on developed system}}{18}{22} \\
  
 \ganttbar[bar/.append style={fill=groupblue}] {\textbf{Analysis and Comparison of results}}{20}{23} \\
 
 \ganttbar[bar/.append style={fill=groupblue}] {\textbf{Thesis Writing}}{9}{28} \\
 
 \ganttbar[bar/.append style={fill=barblue}]{Restructure, correction and adaption of work from PIC2 to Master Thesis report}{9}{28} \\
 \ganttbar[bar/.append style={fill=barblue}]{Reporting about the defined ISA Extention and implemented architecture on the CVA6}{17}{28} \\
 \ganttbar[bar/.append style={fill=barblue}]{Writing of comparisons, analysis and conclusions}{23}{28} 
  
\end{ganttchart}
}
 \caption{Gantt chart of the proposed work plan}
\label{fig:ganttchart}
\end{figure}
\end{landscape}